\section{Conclusion}
\label{section:unsorted/conclusion}

Maintaining the current rate of progress in drug development is a challenging problem.
This is despite advancements in the rate at which new protein structures are being determined.
Perhaps more important, however, is improving the efficiency of discovering successful new drugs, as failures along the drug discovery pathway represent serious economic and health-related costs both to industry and patients.
Computer assisted drug development has the potential not only to maintain, but to improve this progress by helping focus research on more likely drug molecules and allowing efficient screening of large libraries of small molecules. 

The work presented here has focused on development in the Protein Local Optimization Program (PLOP), which has in the past been used as a tool to predict side chain and loop conformations.
We have focused on adding additional support for small molecules and mutation scanning within PLOP and demonstrate a significant algorithmic improvement in computation of implicit solvent models.
The latter improvement greatly enhances performance of PLOP on large systems, allowing more detailed sampling to be performed more efficiently.
We have also demonstrated successful side chain conformation predictions in protein-protein interfaces in our alanine scanning experiments, although predicted binding affinities were not well correlated with experimental data, an area for possible further investigation.
Monte Carlo minimization was successfully implemented and applied to refine docked structures to transition state like complexes, which allowed for successful prediction of sites of metabolism.
This, especially, represents a utilization of PLOP that is highly relevant to the medical community.
Finally, we have made additional progress on a variety of supporting tools that have greatly helped development on the PLOP project in a number of areas.
Taken together, this work has moved the PLOP project forward a great deal, making it more efficient, accurate, and reliable.
